### 2020-10-26 Discussion/Presentation.tex
PUB

\documentclass{beamer}

\usepackage[utf8]{inputenc}

\usetheme{Madrid}
\usecolortheme{beaver}
%Information to be included in the title page:

\title[COVID-19 - Unsupervised Learning]{A statistical analysis on factors that contributed/slowed down the spread of COVID-19}

\author{Marzio De Corato}

\subtitle{Project for the exam: Machine learning, statistical
learning, deep learning and artificial intelligence - Unsupervised Learning}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Overview}
\textbf{FINAL GOAL}: Survey the statistical correlations between the COVID-19 cases/deaths and a selected set of attributes for different clusters via the principal component analysis

\end{frame}

\begin{frame}

\begin{columns}
\frametitle{Cluster features}
\column{0.3\textwidth}
\textbf{Provices (IT)}
\begin{itemize}
\item Unemployment '19
\item Private Transport '12 
\item Air Quality '19
\item Public Transport '12
\item Density '19
\item Mean income '19
\end{itemize}
\column{0.3\textwidth}
\textbf{Regions (IT)}
\begin{itemize}
\item Mortality
\item Pop. for GP
\item Mean Income '19
\item Number of Visits for Pop. '17
\item Public Transport 
\item Density '19
\item Mean income '19
\item Tests 
\item LEA '17
\item Number of visit '17
\item Public structures '17
\end{itemize}
\column{0.3\textwidth}
\textbf{Countries}
\begin{itemize}
\item PM$_{2.5}$ '16
\item Traffic Mortality '16
\item Pollution Mortality '16
\item GDP pro-capita '19
\item Health expenditure '17
\item UHC '17
\end{itemize}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Datasets features }

\textbf{TEMPORAL INTERVAL}: Provinces up to 24/08/2020; Regions up to 24/08/2020; Countries up to 27/08/2020 
\newline
\newline
\textbf{SOURCES}: Protezione Civile, Istituto Italiano di Statistica, Ministro delle Finanze, Ministro della Salute, World Health Organization and World Bank

\end{frame}

\end{document}